Computational methods in landscape ecology

### 2024-08-29 Sections/03-connectivity.qmd
Starting to work on connectivity

<!-- Person responsible: -->

# Connectivity / Movement / Dispersal

<!-- 
https://link.springer.com/article/10.1007/s10980-023-01741-8
https://link.springer.com/article/10.1007/s10980-022-01504-x?fromPaywallRec=false
https://link.springer.com/article/10.1007/s10980-019-00935-3?fromPaywallRec=false
https://link.springer.com/article/10.1038/s41598-022-20370-w?fromPaywallRec=true
-->

>>>>

<!-- Person responsible: Hesselbarth, ??? -->

# Connectivity

Connectivity describes how movement among resources patches is facilitated or hindered by landscape features such as types, amounts, and arrangement of LULC [@Taylor1993; @Taylor2006].
While structural connectivity focuses on physical and spatial relationships among resource patches, functional connectivity also includes the movement or flow of individuals and populations through the landscape [@Taylor2006].
Connectivity underpins ecosystem functionality and maintains biodiversity, and plays a large role in many conservation actions [@CorreaAyram2016; @Hilty2020].
Thus, in the past multiple approaches how to measure connectivity were developed [@Simpkins2018; @Keeley2021].

## Landscape metrics

- Many traditional lsm measure connectivity, e.g, ENN
- Recently, also focus in within-patch connectivity (https://doi.org/10.1007/s10980-019-00881-0)

## Resistance surfaces

https://doi.org/10.1111/2041-210X.12750
https://doi.org/10.1007/s10980-019-00789-9
https://doi.org/10.1007/s10980-022-01504-x
https://doi.org/10.1007/s10980-022-01469-x

## Circuit theory-based

https://circuitscape.org
https://doi.org/10.1111/2041-210X.12689
https://doi.org/10.1016/j.ecolind.2016.03.032

##  Least-cost paths (LCPs)

https://doi.org/10.1016/S0169-2046(02)00242-6
https://doi.org/10.1007/s40823-016-0006-9

## Graph theory

https://doi.org/10.1007/s10980-020-00998-7
https://doi.org/10.1007/s10980-020-01164-9

## Various

https://doi.org/10.1890/14-1690.1
https://doi.org/10.1016/j.ecolind.2016.03.032
https://doi.org/10.1007/s10980-019-00935-3
https://doi.org/10.1126/science.abg1780
